\documentclass[slides]{pgnotes}

\title{Migration}

\begin{document}

\maketitle

\tableofcontents

\section{Cloud utilisation}\label{cloud-utilisation}

\subsection{System}\label{system}

Systems of interest can be quite varied. Remember that not all systems
are 3-tier web or MS-AD environments - there is a huge world out there!
Some of the most interesting systems are the less usual!

\subsection{Provision models}\label{provision-models}

\begin{description}
\item[On-site]
using data centre environment (of varying quality).
\item[Co-lo]
in a dedicated data centre facility.
\item[Cloud]
using 1 or more cloud providers. (Cloud provider as per Nist definition
of cloud computing)
\end{description}

\subsection{Pull factors}\label{pull-factors}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Replacement of capital expenditure by operational expenditure.
\item
  Avoiding sunk hardware costs and upgrade cycles.
\end{enumerate}

\subsection{Push factors}\label{push-factors}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Cost of maintaining on-site data centre environments and their
  supporting infrastructure (e.g.~power, cooling, monitoring,
  networking).
\item
  Demand for space / reduced rental costs
\end{enumerate}

\section{Candidate systems}\label{candidate-systems}

Important to consider what can be moved vs what should be moved.

\subsection{System decomposition}\label{system-decomposition}

System needs to be broken down into its constituent components. It is
best to consider cloud migration as a possibility for individual
components of an application, rather than the system itself as a unit.

\subsection{Strategies}\label{strategies}

We assume our \emph{source environment} is the setup we are currently
using. AWS have identified
\href{https://docs.aws.amazon.com/prescriptive-guidance/latest/migration-replatforming-cots-applications/apg-gloss.html}{7
Rs} describing different migration strategies:

\begin{description}
\item[Retire:]
decommission / remove applications that are no longer needed in your
source environment.
\item[Retain:]
source environment. Candidates:

\begin{itemize}
\item
  Applications requiring major refactoring, and you want to postpone
  that work until a later time.
\item
  May be legacy applications without business justification for
  migrating them.
\end{itemize}
\item[Repurchase / replace (drop and shop)]
switch to a different product (move from custom / COTS to SaaS model).
Example: Migrate your customer relationship management (CRM) system to
Salesforce.com.
\item[Rehost (lift and shift)]
move application cloud without making changes. Example: Migrate
on-premises database / server stack from onsite server to an EC2
instance in the AWS Cloud.
\item[Relocate (hypervisor-level rehost)]
extension of the rehost model where an already virtualised environment
is relocated to the cloud.
\item[Replatform (lift and reshape):]
move an application to the cloud, replacing some components with
cloud-based equivalents. Example: Migrate your on-premises Oracle
database to Amazon Relational Database Service (Amazon RDS) for Oracle
in the AWS Cloud.
\item[Refactor/re-architect:]
modify architecture to use cloud-native capabilities. Example: Migrate
your on-premises Oracle database to the Amazon Aurora
PostgreSQL-Compatible Edition.
\end{description}

Moving any particular system will may not be a single choice.

\subsection{Contra-indications}\label{contra-indications}

Some components present extreme difficulties (not insurmountable)
migrating to cloud environments:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Non-PC/ARM based systems.
\item
  \textbf{Mainframe} systems (e.g.~IBM z/OS) and so-called
  \emph{midrange} (e.g.~IBM i) systems.
\item
  \textbf{Legacy PC operating systems} (e.g.~OS/2, Xenix, PICK, DOS),
  particularly early versions without TCP/IP support.
\item
  Systems requiring \textbf{access to hardware} (e.g.~serial, parallel,
  specific interface cards)
\item
  Systems that \textbf{cannot function off-line} in the absense of
  connectivity (e.g.~some business-critical systems)
\end{enumerate}

May still be possible to move certain components of systems containing
the above to cloud environments if the system can be clearly decomposed
into interconnecting parts.

\subsection{Onsite connectivity}\label{onsite-connectivity}

Moving

Possible to connect a VPC to an onsite LAN using a VPN gateway.

Some co-lo data centres offer Amazon Direct Connect (not discussed in
detail here).

\end{document}

